---
title: "Mixed ANOVA + Pairwise Comparison (CL data)"
date: "6/3/2021"
output:
  html_document:
    toc: true
    df_print: paged
    theme: journal
    code_folding: hide
    
knit: 
    (function(input, ...) {
        rmarkdown::render(input, output_file = "RAnova_main.html")
    })
---

# Description

**Source of motivation: [Mixed ANOVA in R](https://www.datanovia.com/en/lessons/mixed-anova-in-r)**

## Assumptions Check

-   **Outliers**. Values above `Q3 + 1.5xIQR` or below `Q1 - 1.5xIQR` are considered as outliers. Values above `Q3 + 3xIQR` or below `Q1 - 3xIQR` are considered as extreme points (or extreme outliers). Observations that are considered as extreme points are marked red.

-   **Normality Assumptions**. Plot shows Q-Q plot for subsets of data. If points lie on the line, we can assume that data is fairly symmetrical (and that's what we want).

-   **Homogeneity of Variance Assumption**. Performed using [Levene's test](https://en.wikipedia.org/wiki/Levene%27s_test).

    -   $H_0$: the groups have equal population variances
    -   $H_A$: the groups don't have equal population variances
    -   Observations with p-value \< 0.05 are marked red, since they may violate the ANOVA results.

## Mixed ANOVA

"*In statistics, a mixed-design analysis of variance model, also known as a split-plot ANOVA, is used to test for differences between two or more independent groups whilst subjecting participants to repeated measures.*" ([Wikipedia](https://en.wikipedia.org/wiki/Mixed-design_analysis_of_variance))

-   within subjects group (with repeated measures): `Measurement`
-   between subjects group: `Group`

## Post-hoc Tests

-   **Significant two-way interaction**. Can be used if an interaction term in Mixed ANOVA results showed significance. In other words, impact of `Group` on the outcome variable (`PeakAmp`, `PeakLat`, `RMS`) depends on the `Measurement` (and vice versa)

    -   **Pairwise comparisons between `Group`**. Performs [Welch's t-tests](https://en.wikipedia.org/wiki/Welch%27s_t-test) (for unequal variances) for comparing `Group` for each `Measurement`

    -   **Pairwise comparisons between `Measurement`**. Performs **paired** two-sample t-test for comparing `Measurement` for each `Group`.

-   **Non-significant two-way interaction**. Shows the **main effect** for `Group` and `Measurement`. Used if there was non-significant two-way interaction.

    For each table observations with adjusted p-value (using Holm–Bonferroni method) which is less then 0.05 are marked green.

```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(ggpubr)
library(rstatix)
library(knitr)
library(kableExtra)
```

```{r message=FALSE, warning=FALSE}
data_type <- "CLST"
raw_data <- read_csv(paste0("../../Data_new/AVRECPeak", data_type,".csv"))

layers <- unique(raw_data$Layer)
stimfreq <- unique(raw_data$ClickFreq)
metrics <- c("PeakAmp", "PeakLat", "RMS", "logRMS")
stimfreq <- c(5, 10)


red_color <- "#ffb3b3"
green_color <- "#d6f5d6"

# values for testing
# metrics <- c("RMS", "logRMS")
freq <- 5
layer <- "All"
metric <- "logRMS"
```

# Results {.tabset .tabset-fade .tabset-pills}

```{r message=FALSE, warning=FALSE, results='asis'}
for (freq in stimfreq){

  cat("##", freq, "Hz Frequency {.tabset .tabset-fade .tabset-pills} \n")
  
  for (layer in layers){
  # for (layer in c("All", "I_II")){
    cat("###", layer, " Layer {.tabset .tabset-fade .tabset-pills}\n")
    
    model_data <- raw_data %>% 
      filter(
        ClickFreq == freq,
        Layer == layer,
        OrderofClick == 1,
        grepl('_1', Measurement),
        PeakAmp > 0) %>% 
      mutate(
        logRMS = log(RMS),
        Measurement = case_when(
          grepl('pre', Measurement) ~ "preLaser",
          TRUE ~ "postLaser"
        )
      ) %>% 
      select(-c(Layer, ClickFreq, OrderofClick))
  
    # get average by Trial
    if(grepl("ST", data_type)){
      model_data <- model_data %>%
        group_by(Group, Animal, Measurement) %>%
        summarize(
          PeakAmp = mean(PeakAmp, na.rm = TRUE),
          PeakLat = mean(PeakLat, na.rm = TRUE),
          RMS = mean(RMS, na.rm = TRUE),
          logRMS = mean(logRMS, na.rm = TRUE)) %>%
        ungroup() %>% 
        convert_as_factor(Group, Animal, Measurement) 
    }
    
    # if(grepl("ST", data_type)){
    #   model_data <- model_data %>% 
    #     mutate(Animal = paste0(Animal, "_", TrialNumber)) %>% 
    #     select(-TrialNumber) %>% 
    #     ungroup()
    # }
  
    # exclude_animals <- model_data %>%
    #   group_by(Animal, Layer) %>%
    #   summarise(n = n_distinct(Measurement)) %>%
    #   filter(n < 2) %>%
    #   pull(Animal)
    # 
    # model_data <- model_data %>% 
    #   filter(!(Animal %in% exclude_animals)) %>%
    #   select(Group, Animal, Measurement, Layer, PeakAmp, PeakLat, RMS, logRMS) %>% 
    #   convert_as_factor(Group, Animal, Measurement, Layer) 
    
      for (metric in metrics){
      # for (metric in c("PeakLat")){

        cat("####", metric, "{.tabset .tabset-fade .tabset-pills}\n")
        
        cat("##### Assumptions Check\n")
        
        cat("<b>===OUTLIERS===</b>\n\n")
        
        df <- model_data %>%
          group_by(Measurement, Group) %>%
          identify_outliers(eval(metric))
        
        if(nrow(df) > 0){
          df %>% 
          kable(escape = F, digits = 6) %>%
          kable_classic_2(full_width = F) %>% 
          kable_styling() %>% 
          row_spec(
            which(df$is.extreme == TRUE), bold = T, 
            background = red_color) %>% 
          print()
        } else{
          cat("\n\nNo outliers were detected.\n\n")
        }
        
        cat("\n\n<b>===NORMALITY ASSUMPTION===</b>\n\n")
        
        p <- ggqqplot(model_data, metric, ggtheme = theme_bw()) +
          facet_grid(Group ~ Measurement, labeller = "label_both")
        plot(p)

        cat("\n\n<b>===HOMOGENEITY OF VARIANCE ASSUMPTION===</b>\n\n")
        
        df <- model_data %>%
          group_by(Measurement) %>%
          levene_test(formula = as.formula(paste0(metric, "~ Group"))) 
        
        df %>% 
          kable(escape = F, digits = 3) %>%
          kable_classic_2(full_width = F) %>% 
          kable_styling() %>% 
          row_spec(
            which(df$p <= 0.05), bold = T, 
            background = red_color) %>% 
          print()
        
        cat('\n\n')

        cat("##### Mixed ANOVA\n")

        # Two-way mixed ANOVA test
        res.aov <- anova_test(
          data = model_data, dv = eval(metric), wid = Animal,
          between = Group, within = c(Measurement))
        
        df <- get_anova_table(res.aov) %>%
          data.frame()
        
        df %>% 
          kable(escape = F,
                caption = "Mixed ANOVA (Type III test)") %>%
          kable_classic_2(full_width = F) %>% 
          kable_styling() %>% 
          row_spec(
            which(df$p <= 0.05), bold = T, 
            background = green_color) %>% 
          print()

        cat('\n\n')

        cat("##### Post-hoc Tests\n")

        cat("\n\n<b>===SIGNIFICANT TWO-WAY INTERACTION===</b>\n\n")
        
        # cat("\n\n\n<b>=EFFECT OF GROUP ON MEASUREMENT=</b>\n\n")
        
        # one.way <- model_data %>%
        #   group_by(Group) %>%
        #   anova_test(dv = eval(metric), wid = Animal, within = c(Measurement)) %>% 
        #   get_anova_table() %>%
        #   adjust_pvalue(method = "holm")
  
        # one.way <- model_data %>%
        #   group_by(Measurement) %>%
        #   anova_test(dv = eval(metric), wid = Animal, between = Group) %>%
        #   get_anova_table() %>%
        #   adjust_pvalue(method = "bonferroni")
        
        # df <- one.way %>% 
        #   data.frame() %>% 
        #   select(-F)
        # df %>% 
        #   kable(escape = F, digits = 4) %>%
        #   kable_classic_2(full_width = F) %>% 
        #   kable_styling() %>% 
        #   row_spec(
        #     which(df$p.adj <= 0.05), bold = T, 
        #     background = green_color) %>% 
        #   print()
        
        cat("\n\n\n<b>=PAIRWISE COMPARISONS BETWEEN GROUP LEVELS=</b>\n\n")
        
        pwc <- model_data %>%
          group_by(Measurement) %>%
          pairwise_t_test(
            as.formula(paste0(metric, "~ Group")),
            p.adjust.method = "holm",
            pool.sd = FALSE,
            var.equal = FALSE)

        eff_size <- model_data %>%
          group_by(Measurement) %>%
          cohens_d(
            as.formula(paste0(metric, "~ Group")),
            var.equal = FALSE)
        
        df <- pwc %>% 
          left_join(eff_size) %>% 
          data.frame() 
        
        df %>% 
          kable(escape = F, digits = 3) %>%
          kable_classic_2(full_width = F) %>% 
          kable_styling() %>% 
          row_spec(
            which(df$p.adj <= 0.05), bold = T, 
            background = green_color) %>% 
          print()
        
        # cat("\n\n\n<b>=EFFECT OF MEASUREMENT AT EACH LEVEL OF GROUP=</b>\n\n")
        # 
        # one.way2 <- model_data %>%
        #   group_by(Measurement) %>%
        #   anova_test(dv = eval(metric), wid = Animal, between = Group) %>%
        #   get_anova_table() %>%
        #   adjust_pvalue(method = "holm")
        # 
        # df <- one.way2 %>%
        #   data.frame()
        # df %>% 
        #   kable(escape = F) %>%
        #   kable_classic_2(full_width = F) %>% 
        #   kable_styling() %>% 
        #   row_spec(
        #     which(df$p.adj <= 0.05), bold = T, 
        #     background = green_color) %>% 
        #   print()
        
        cat("\n\n\n<b>=PAIRWISE COMPARISONS BETWEEN MEASUREMENTS AT EACH GROUP LEVELS=</b>\n\n")
        
        # exclude_animals <- model_data %>%
        #   group_by(Animal) %>%
        #   summarise(n = n_distinct(Measurement)) %>%
        #   filter(n < 2) %>%
        #   pull(Animal)
        
        pwc2 <- model_data %>%
          # filter(!(Animal %in% exclude_animals)) %>%
          group_by(Group) %>%
          pairwise_t_test(
            as.formula(paste0(metric, "~ Measurement")), 
            paired = TRUE, 
            pool.sd = FALSE,
            var.equal = FALSE,
            p.adjust.method = "holm"
          )
        
        eff_size2 <- model_data %>%
          # filter(!(Animal %in% exclude_animals)) %>%
          group_by(Group) %>%
          cohens_d(
            as.formula(paste0(metric, "~ Measurement")), 
            paired = TRUE, 
            var.equal = FALSE)

        df <- pwc2 %>% 
          left_join(eff_size2) %>% 
          data.frame()
        df %>% 
          kable(escape = F, digits = 3) %>%
          kable_classic_2(full_width = F) %>% 
          kable_styling() %>% 
          row_spec(
            which(df$p.adj <= 0.05), bold = T, 
            background = green_color) %>% 
          print()
        
        cat("\n\n\n<b>===NON-SIGNIFICANT TWO-WAY INTERACTION===</b>\n\n")
        cat("\n\n\n<b>=MAIN EFFECT OF MEASUREMENT=</b>\n\n")
        
        df <- model_data %>%
          # filter(!(Animal %in% exclude_animals)) %>%
          pairwise_t_test(
            as.formula(paste0(metric, "~ Measurement")), 
            paired = TRUE, 
            p.adjust.method = "holm"
          ) %>% 
          data.frame() 
    
        df %>% 
          kable(escape = F) %>%
          kable_classic_2(full_width = F) %>% 
          kable_styling() %>% 
          row_spec(
            which(df$p.adj <= 0.05), bold = T, 
            background = green_color) %>% 
          print()
        
        cat("\n\n\n<b>=MAIN EFFECT OF GROUP=</b>\n\n")
        
        df <- model_data %>%
          pairwise_t_test(
            as.formula(paste0(metric, "~ Group")), 
            p.adjust.method = "holm"
          ) %>% 
          data.frame() 
        
        df %>% 
          kable(escape = F) %>%
          kable_classic_2(full_width = F) %>% 
          kable_styling() %>% 
          row_spec(
            which(df$p.adj <= 0.05), bold = T, 
            background = green_color) %>% 
          print()
        
        cat('\n\n')
        
        cat("##### Plots\n")
        
        pwc <- pwc %>% add_xy_position(x = "Measurement")
        pwc2 <- pwc2 %>% add_xy_position(x = "Group")
        
        box_plot <- ggboxplot(
          model_data, x = "Measurement", y = metric,
          color = "Group", palette = "jco",
          add = "jitter"
        )
        
        box_plot2 <- ggboxplot(
          model_data, x = "Group", y = metric, 
          color = "Measurement", palette = "jco",
          add = "jitter"
        )
        
        p <- box_plot + 
          stat_pvalue_manual(pwc, tip.length = 0.01, hide.ns = TRUE) +
          labs(
            title = paste0(metric, " for ", layer, " Layer, ", freq, "Hz and 1st Click"),
            subtitle = get_test_label(res.aov, detailed = TRUE),
            caption = get_pwc_label(pwc)
          )
        
        plot(p)
        
        cat('\n\n')
        
        p <- box_plot2 + 
          stat_pvalue_manual(pwc2, tip.length = 0.01, hide.ns = TRUE) +
          labs(
            title = paste0(metric, " for ", layer, " Layer, ", freq, "Hz and 1st Click"),
            subtitle = get_test_label(res.aov, detailed = TRUE),
            caption = get_pwc_label(pwc2)
          )
        
        plot(p)

        cat('\n\n')
      
      cat('\n\n')
    }
    
    cat('\n\n')
  }
  
  cat('\n\n')
}
```

```{r}
anova_test(
          data = model_data, dv = logRMS, wid = Animal,
          between = Group, within = Measurement, type = "II")
```

```{r}
library(nlme)

test <- lme(fixed = logRMS ~ Group*Measurement,
            random = ~1|Animal/Measurement,
            data = model_data)
anova(test, type = "sequential")
```

```{r}
library(gridExtra)
p1 <- plot(test, resid(.) ~ fitted(.), abline=0, main = "All Data")
p2 <- plot(test, resid(.) ~ fitted(.)|Group, abline=0, main = "By Animal")

grid.arrange(p1, p2, nrow = 1, widths = c(1.5,2))
```


```{r eval=FALSE, include=FALSE}
library(emmeans)
emmip(test, formula = Group ~ Measurement) +
        theme_minimal()

```

```{r}
pairs(emmeans(test, ~ Group | Measurement))
```

```{r}
model_data %>%
  # filter(!(Animal %in% exclude_animals)) %>%
  group_by(Measurement) %>%
  pairwise_t_test(
    as.formula(paste0(metric, "~ Group"))
  )
```


```{r}
model_data %>% 
  group_by(Group, Measurement) %>% 
  summarise(mean(logRMS))
```

```{r}

```


```{r}
pairs(lsmeans(test, ~ Group | Measurement))
```

```{r}
model_data <- raw_data %>% 
      filter(
        ClickFreq == freq,
        Layer == "V",
        OrderofClick == 1,
        grepl('_1', Measurement),
        PeakAmp > 0) %>% 
      mutate(
        logRMS = log(RMS),
        Measurement = case_when(
          grepl('pre', Measurement) ~ "preLaser",
          TRUE ~ "postLaser"
        )
      ) %>% 
      select(-c(Layer, ClickFreq, OrderofClick))
  
    # get average by Trial
    if(grepl("ST", data_type)){
      model_data <- model_data %>%
        group_by(Group, Animal, Measurement) %>%
        summarize(
          PeakAmp = mean(PeakAmp, na.rm = TRUE),
          PeakLat = mean(PeakLat, na.rm = TRUE),
          RMS = mean(RMS, na.rm = TRUE),
          logRMS = mean(logRMS, na.rm = TRUE)) %>%
        ungroup() %>% 
        convert_as_factor(Group, Animal, Measurement) 
    }

ggboxplot(
          model_data, x = "Group", y = "RMS",
          color = "Measurement", palette = "jco",
          add = "mean", ylim = c(0,0.005)
        )
```

```{r}
model_data <- raw_data %>% 
      filter(
        ClickFreq == freq,
        Layer == "V",
        OrderofClick == 1,
        grepl('_1', Measurement),
        PeakAmp > 0) %>% 
      mutate(
        logRMS = log(RMS),
        Measurement = case_when(
          grepl('pre', Measurement) ~ "preLaser",
          TRUE ~ "postLaser"
        )
      ) %>% 
      select(-c(Layer, ClickFreq, OrderofClick))
  

ggboxplot(
          model_data, x = "Group", y = "RMS",
          color = "Measurement", palette = "jco",
          add = "mean", ylim = c(0,0.005)
        )
```